Heritability

### 2022-07-05 Manuscript/manuscript.Rmd
combine all edits to incorporate the figures and analyses into the manuscript

# CH4 models
load(file = '../Output/response_model.RData')
load(file = '../Output/herit_model.RData')

# 16S models
load(file = '../Output/richness_models.RData')
load(file = '../Output/phyloseq.RData')
load(file = '../Output/beta_model.RData')
load(file = '../Output/da_corncob.RData')

@@

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`). In addition, the
response to selection varied with treatment 
`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

@@

Sequencing of 16S rRNA amplicons revealed that there were `r ntaxa(ps)` unique
ASVs across all `r nsamples(ps)` jars sampled.
Richness of ASVs decreased among all of the jars between Passage 2 and 5
(Richness difference between Passage 5 and Passage 2 = `r myround(bt_passage$table[2, 1], 1)`, SE =
`r myround(bt_passage$table[2, 2], 2)`, p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 5 (Difference = `r myround(bt_treat5$table[2, 1], 1)`, SE =
`r myround(bt_treat5$table[2, 2], 2)`, p =
`r myround(bt_treat5$table[2, 3], 3)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across all passages
and treatments.  (slope = `r myround(bt_ch4$table[2, 1], 1)`, SE =
`r myround(bt_ch4$table[2, 2], 2)`, p = `r myround(bt_ch4$table[2, 3], 3)`).

Taxonomic dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment (`r figures("beta", display = "cite")`). Permutational
analysis of variance demonstrated an effect of passage, treatment,
and their interaction on Aitchison dissimilarity (`r tables("adonis", display = "cite")`). 
Passage explained `r myround(100*fit_adonis[1, 4])`\% of the
variation in Aitchison dissimilarity and `r myround(100*fit_adonis[6, 4])`\% in
Jaccard dissimilarity.  Treatment explains `r myround(100*fit_adonis[2, 4])`\%
for Aitchison and `r myround(100*fit_adonis[7, 4])`\% for Jaccard.  Variation
in microbiome composition decreased as a result of passaging in jars (F = 80.5,
p < 0.001), but did not differ between selection treatments (F = 0.40, p =
0.54).

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
Silvanigrellaceae, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
Silvanigrella is the only cultivated member of
Silvanigrellaceae and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, Armatimonadaceae, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for Armatimonadaceae was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

The remaining families were shared between the Positive and Neutral treatment,
but some differed in their relative abundance. To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using a beta-binomial model. We identified `r ntaxa(sig_taxa_treat)`
families that were enriched or depleted in the Positive treatmeant relative to
the Neutral treatment (`r figures("da", display = "cite")`).

Overall, none of the families enriched in the Positive selection treatment
are known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
This class is one of the groups containing methanotrophic families such as
Methylococcaceae. However, gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
The next family
was Puniceicoccaceae, which is a member of the phylum Verrucomicrobia and
contains marine microbes. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. 
An ASV from the order Armatimonadales was also enriched in the Positive treatment.
This order also contains the family Armatimonadaceae, which was
found to be unique to the Neutral treatment. Cytophagaceae was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
Gemmatimonadaceae has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
that contains methanotrophy genes, but this family was depleted in the Positive
selection treatment [@bay2021].

@@

```{r response, fig.width=4, fig.height=3}
ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)

p <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
p
ggsave('fig1.jpg', p)

@@

p <- plot_herit(heritability)
p
ggsave('fig2.jpg', p)

@@

p <- plot_beta(ps)
p
ggsave('fig3.jpg', p)

@@

p <- dt_plot(dt_data)
p
ggsave('fig4.jpg', p)

@@

`r tables("adonis", "PERMANOVA on Aitchison and Jaccard dissimilarities.")`

@@

write.table(fit_adonis, 'table_1.tsv')
fit_adonis %>%
  mutate(metric = c(rep("aitchison", 5), rep("jaccard", 5))) %>% 
  select(metric, term:p.value) %>% 
	flextable()
# kable(col.names = c('Term', 'df', 'SS', 'R2', 'F', 'p'), 
#       digits = c(2,2,2,2,2,3),
## 	format = "Table",
## 	booktabs = T,
# 	caption = "PERMANOVA on Bray-Curtis and Jaccard dissimilarities.",
#   	label = "beta") %>%
#   pack_rows("Bray-Curtis", 1, 5) %>%
#   pack_rows("Jaccard", 6, 10)

>>>>

# CH4 models
load(file = '../Output/response_model.RData')
load(file = '../Output/response_plot.RData')
load(file = '../Output/herit_model.RData')
load(file = '../Output/herit_plot.RData')

# 16S models
load(file = '../Output/richness_models.RData')
load(file = '../Output/phyloseq.RData')
load(file = '../Output/beta_model.RData')
load(file = '../Output/beta_fig.RData')
load(file = '../Output/da_corncob.RData')
load(file = '../Output/da_fig.RData')

@@

## Response to selection on methane oxidation rate

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; `r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

## Heritability of methane oxidation

@@

## Taxonomic richness

Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: chi-squared = 
`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; Pairwise Wilcoxon test: p < 0.001). However, there was no
difference in richness between the Positive and the Neutral treatment in
Passage 2 or 5 (Pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
correlation between richness and CH~4~ oxidation rate across all passages
and treatments (Spearman's rho = `r myround(s[1,1], 1)`, p = `r myround(s[1,3], 1)`).

## Community dissimilarity

Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
variation in Bray-Curtis dissimilarity.  Treatment explained 
`r myround(100*tidy_permanova[2, 4])`\%
of the variation. There was also an interaction
between treatment and passage, which explained 
`r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45).

## Taxa that responded to selection

To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
first looked at which taxa were unique to the Positive and Neutral treatments
aggregated at the family level. Several taxa were present in the Positive
treatment and absent in the Neutral treatment. Most of these families had low
prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
There were two families unique to the Positive treatment with relatively high
prevalence. This included an ASV that was a member of the Bacteroidia Class with
no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
abundance of ~2 reads. The other prevalent family was a member of the
*Silvanigrellaceae*, a newly described family placed in its own order. This family
was present in all 12 samples and had a median abundance of 8 reads.
*Silvanigrella* is the only cultivated member of
*Silvanigrellaceae* and was isolated from a temperate fresh water lake
[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
the rhizosphere of *Phragmites australis* [@tamaki2011].

The remaining families were shared between the Positive and Neutral treatment,
but some differed in their relative abundance. To identify taxa that responded
to selection on soil CH~4~ oxidation rate, we tested the differential relative
abundance of families in the Positive jars relative to the Neutral jars within
Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
families that were enriched or depleted in the Positive treatment relative to
the Neutral treatment (`r figures("da", display = "cite")`).

Overall, none of the families enriched in the Positive selection treatment
are known methanotrophs. Several taxa identified had a 
higher taxonomic designation that contains methanotrophs, for example, the
Gammaproteobacteria class had a large effect size.
This class is one of the groups containing methanotrophic families such as
*Methylococcaceae.* However, Gammaproteobacteria is among the most diverse 
groups in the Prokaryotes, so this is not strong evidence for a selection response
by methanotrophs [@garrity2005].
The next family
was *Puniceicoccaceae*, which is a member of the phylum Verrucomicrobia and
contains marine microbes. The Verrucomicrobia is a diverse group that contain
known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. 
An ASV from the order Armatimonadales was also enriched in the Positive treatment.
This order also contains the family *Armatimonadaceae*, which was
found to be unique to the Neutral treatment. *Cytophagaceae* was also enriched in
the Positive treatment and contains a number of mainly aerobic heterotrophs
that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
*Gemmatimonadaceae* has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
that contains methanotrophy genes, but this family was depleted in the Positive
selection treatment [@bay2021].

@@

```{r response, fig.width=4, fig.height=3}

p_response

@@

p_herit

@@

p_beta

@@

p_log10

@@

`r tables("adonis", "PERMANOVA on Bray-Curtis dissimilarities.")`

@@

tidy_permanova %>%
	flextable()

### 2022-07-05 R/04-heritability.Rmd
combine all edits to incorporate the figures and analyses into the manuscript

@@

```{r initialize_variables}
source('functions.R')
```

@@

plot_herit(heritability)

@@

```{r save_heritability}
save(herit_model, file = '../Output/herit_model.RData')
```

>>>>

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
knitr::opts_knit$set(root.dir=normalizePath('../'))

@@

```{r initialize_variables}
source('R/functions.R')
```

@@

p_herit <- plot_herit(heritability)

p_herit

save(p_herit, file = 'Output/herit_plot.RData')

ggsave('Manuscript/fig2.jpg', p_herit, width = 4, height = 3)

@@

```{r save_heritability}
save(herit_model, file = 'Output/herit_model.RData')
```
